Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-08-01 manuscript/manuscript.tex
Continue writing manuscript. Added figs: linear density tests

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/spherical-shell.pdf}
\caption{
Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
The computation point $Q$ is located in the $z$ axe at a distance $r$ from the centre of the shell.
For practical purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
}
\label{fig:spherical-shell}
\end{figure}

@@

In this way, the numerical approximation efficiently improves its accuracy and
gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g. less than 10cm), which does not modifies significantly the desired gravitational field.

On the other hand, the accuracy is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
In order to solve this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.

@@

>>>>

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{figures/spherical-shell.pdf}
\caption{
Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
The computation point $Q$ is located in the $z$ axe at a distance $r$ from the centre of the shell.
For practical purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
}
\label{fig:spherical-shell}
\end{figure}

@@

In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided and therefore both the accuracy of the algorithm and its computation time.
In this way, the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g. less than 10cm), which does not modifies significantly the desired gravitational field.

On the other hand, the accuracy of the method is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
In order to solve this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.

@@

\section{Determination of distance-size ratio}

In order to test the new variable density tesseroid code and to estimate the error of the approximation we have compared the numerical model with the analytical solutions for a spherical shell with variable density for two special cases: a linear and an exponential density function.
This error estimations are also useful in the determination of the value for the distance-size ratio $D$, such the difference between the numerical model and the analytical solutions are below an acceptable threshold.

We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the $g_{zz}$ component of the Marussi tensor on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
More information about this grids can be found in the Table \ref{tab:grids}.

\todo{explicar pq los otros campos no los testeamos}

We computed the three gravity fields for different values of $D$, ranging from 1 to 10 with a step of 0.5 on every point of these four grids. Then, we calculated the maximum difference between these results and the value of the analytical solution at the same height.

Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude. That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.

Finally, we set the acceptable value of $D$ at which the error of the numerical approximation is lower than 10$^{-1}\%$.

\begin{table}
\caption{Description of the synthetic grids where the comparison of the numerical model against the analytical solutions for a spherical shell was done. This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height.}
\label{tab:grids}
\begin{tabular}{lcccc}
    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
\end{tabular}
\end{table}

\subsection{Linear Density}

As a simple case, we proposed a linear density function such as the spherical shell has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:

\begin{equation}
    \rho(r') = 
    -\frac{3300 - 2670}{35000} r' + 
    \left( \frac{3300 - 2670}{35000} R + 2670 \right)
\end{equation}

\noindent where $R$ = 6378.137 km is the mean Earth radius.

The tests were performed on a thin and a thick spherical shell of 1 km and 35 km of thickness, respectively.
The results for both cases can be seen in Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.

We can see that for a linear density, the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 10$^{-1}\%$.
On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.
We will keep the more conservative result of $D=8$ in case of computing the $g_{zz}$ tensor component in case of a linear density tesseroid.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dlinear-thin-differences.pdf}
\caption{}
\label{fig:D-linear-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/Dlinear-thick-differences.pdf}
\caption{}
\label{fig:D-linear-thick}
\end{figure}

\subsection{Exponential Density}
